### 2016-02-07 figures/simlex-ppmi-best.tex
SPMI story

\begin{figure*}
  \centering
    \includegraphics[width=\textwidth]{supplement/figures/simlex-ppmi-best.pdf}

  \caption{\textbf{SPMI best configurations.}}
  \label{fig:simlex-ppmi-best}
\end{figure*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: