### 2016-02-07 figures/simlex-ppmi-interaction.tex

\begin{figure*}
  \centering
  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{supplement/figures/simlex-ppmi-cds.pdf}
    \caption{\textbf{Context distribution smoothing.}}
    \label{fig:simlex-ppmi-interaction-cds}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{supplement/figures/simlex-ppmi-neg.pdf}
    \caption{\textbf{Shifting.}}
    \label{fig:simlex-ppmi-interaction-neg}
  \end{subfigure}

  \caption{\textbf{SPMI.}}
  \label{fig:simlex-ppmi-interaction}
\end{figure*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: