### 2016-02-08 figures/simlex-ppmi-best.tex
Complete the section on lexical tasks.

\begin{figure*}
  \centering
  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{supplement/figures/SimLex999-ppmi-best.pdf}
    \caption{SimLex-999}
    \label{fig:simlex-ppmi-best-simlex}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{supplement/figures/men-ppmi-best.pdf}
    \caption{MEN}
    \label{fig:simlex-ppmi-best-men}
  \end{subfigure}

  \caption{\textbf{SPMI best configurations.} Black line shows the best results reported by \protect\newcite{TACL570}.}
  \label{fig:simlex-ppmi-best}
\end{figure*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: