### 2016-02-16 figures/interaction-cds.tex

\begin{figure*}
  \centering
    \includegraphics[width=\textwidth]{supplement/figures/SimLex999-interaction-cds}
    \caption{\textbf{Context distribution smoothing.}}
    \label{fig:interaction-cds}
\end{figure*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: