### 2016-02-27 figures/interaction-cds.tex
Update

\begin{figure*}
  \centering
    \centerline{\includegraphics[width=1.11\textwidth]{supplement/figures/SimLex999-interaction-cds}}
    \caption{\textbf{The effect of context probability estimation and smoothing on SimLex-999.}}
    \label{fig:interaction-cds}
\end{figure*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: