### 2016-02-27 figures/interaction-neg.tex

\begin{figure*}
  \centering
    \centerline{\includegraphics[width=1.11\textwidth]{supplement/figures/SimLex999-interaction-neg}}
    \caption{\textbf{The behaviour of shifted PMI on SimLex-999.}}
    \label{fig:interaction-neg}
\end{figure*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: