### 2016-06-07 figures/interaction-cds.tex
Mentoring comments

\begin{figure*}
  \centering
    \centerline{\includegraphics[width=1.11\textwidth]{supplement/figures/SimLex999-interaction-cds}}
    \caption{{\bf Effect of PMI variant (\texttt{discr}), smoothing (\texttt{cds}) and frequency weighting (\texttt{freq}) on SimLex-999.} Error bars correspond to a 95\% confidence interval as the value is estimated by aggregating over omitted parameters: \texttt{neg} and similarity.}
    \label{fig:interaction-cds}
\end{figure*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: